Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-21 paper/paper.tex
Ewa's comments

\keywords{Scientific workflows, DAG scheduling, simulation}

@@

We assume that we have runtime estimates for each task in the workflow based on
either a performance model of the application, or historical data that can be
mined.

@@

To evaluate and compare our algorithms, we developed a cloud workflow simulator based on
CloudSim~\cite{Calheiros11}. Our simulation model consists of Cloud, VM and WorkflowEngine 
entities. The Cloud entity is responsible for staring and terminating VM entities using 
an API similar to Amazon EC2. VM entities simulate the execution of individual tasks,
and the WorkflowEngine entity manages the scheduling of tasks and the provisioning of
VMs. We assume that the VMs have a single core and execute tasks sequentially. Although 
CloudSim provides a more advanced infrastructure model which includes time-sharing and 
space-sharing policies, and power consumption in a datacenter, we do not use these 
features, since we are interested mainly in the execution of tasks on VMs and high-level 
workflow scheduling and provisioning. The simulator reads workflows in a modified version of
the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005} that was
created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.

@@

We have prepared test scenarios to evaluate the proposed algorithms under a 
range of parameters that enables us to observe the interesting characteristics 
of the different algorithms for smaller and larger constraints. For a given 
ensemble we can expect, in general, that the amount of work done (in terms 
of workflows completed) depends mainly on the budget constraint, which 
determines the number of machine-hours that can be provisioned. The deadline, 
although important in terms of scheduling, is assumed to be greater than the 
critical path of every workflow. The ideal but trivial case, then, is a deadline 
long enough to allow all the workflow tasks to be executed sequentially: this 
eliminates all the overheads of parallel execution and leads to the highest 
resource utilization. More interesting, however, are the cases where the deadline 
is shorter than the serial schedule, since it forces the system to provision 
multiple VMs to execute workflows in parallel.

@@

The number of workflows in an ensemble depends on the user's requirements, but
we assume that ensemble sizes are on the order of between 10 and 100 workflows.
This range is motivated by several reasons: First, such sizes are typical of the
real applications we have examined. For example, the number of geographical 
sites of interest to the users of the CyberShake application is on the order 
of 100. Second, smaller ensembles consisting of just a few workflows can be 
aggregated into a single workflow, so there is no need to treat them as an
ensemble. Similarly, when the number of workflows is larger than 100, and each 
workflow has a large number of tasks, either the constraints are low enough to
prevent many of the workflows from running, or the problem of efficiently 
allocating them to the resources becomes similar to the bag-of-tasks problem, 
which is easier to solve efficiently.

@@

In order to evaluate the algorithms on a standard set of workflows, we used
ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains synthetic
workflows modeled on real applications. We used workflows from five applications 
that were developed for the Pegasus Workflow Management System. We constructed 
ensembles of 100 workflows by sampling collections of workflows of various sizes
(number of tasks) based on a given distribution. Two distributions are considered:

@@

The constant distribution produces ensembles containing of workflows of equal size, 
while the Pareto-like distribution produces ensembles with a small number of larger 
workflows and a large number of smaller workflows, as seen in 
Fig.~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
not-insignificant number of large workflows in the ensemble. We assume that large
workflows are more important to users, so in the Pareto ensembles the workflows are 
assigned priorities according to their size so that the largest workflows are given 
the highest priority. In future work we plan to explore other ensemble types such as
uniform distributions.

@@

In the experiments we assumed that all the VMs are homogeneous, have a
processing speed of 1000 MIPS (million instructions per second), and a price of
\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery 
are given in seconds measured on a 1000 MIPS machine. In this study we do not 
take into account the heterogeneity of the infrastructure since we assume that 
it is always possible to select a VM type that has the best price to performance 
ratio for a given application.

For each application, we selected ranges of constraints (deadline and budget) so
that it is possible to observe the possibly broad result space: from tight 
constraints, where only a small number of workflows can be completed, to 
more liberal constraints where all, or almost all, of the workflows can be 
completed. These parameters vary among the applications. Montage and CyberShake 
ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
20 hours to be completed. The LIGO and SIPHT workflows have more long-running
tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
approximately \$12,000 and 1500 hours to complete a full ensemble.

@@

\begin{itemize}
  \item $N_c$: number of workflows completed,
  \item $s$: user preference score (utility function),
  \item $T_C$: total amount of computing time completed,
  \item $C_{eff}$: effective cost of computing per hour, which includes
  overheads.
\end{itemize}

@@

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-p}}
\end{equation}

@@

The total amount computing time completed, $T_C$, is calculated as the sum 
of the runtimes of all tasks in all completed workflows. Because we assumed 
that the resource cost per hour is equal to \$1, the total computing time 
can be easily compared to the allocated budget $b$. Their ratio: 
$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour.

>>>>

\keywords{Scientific workflows, cloud computing, DAG scheduling, simulation}

@@

We assume that we have runtime estimates for each task in the workflow based on
either a performance model of the application, or historical data that can be
mined. In particular, in this study we use workflows from the synthetic workflow
generator~\cite{Bharathi08}.

@@

To evaluate and compare our algorithms, we developed a cloud workflow simulator based on
CloudSim~\cite{Calheiros11}. Our simulation model consists of Cloud, VM and WorkflowEngine 
entities. The Cloud entity is responsible for staring and terminating VM entities using 
an API similar to Amazon EC2. VM entities simulate the execution of individual tasks,
and the WorkflowEngine entity manages the scheduling of tasks and the provisioning of
VMs. We assume that the VMs have a single core and execute tasks sequentially. Although 
CloudSim provides a more advanced infrastructure model which includes time-sharing and 
space-sharing policies, and power consumption in a datacenter, we do not use these 
features, since we are interested mainly in the execution of tasks on VMs and high-level 
workflow scheduling and provisioning. The simulator reads workflows in a modified version of
the DAX format used by the Pegasus Workflow Management System~\cite{Deelman2005}. The modified format was
created for the synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi08}.

@@

We have prepared test scenarios to evaluate the proposed algorithms under a
range of parameters that enables us to observe the interesting characteristics
of the different algorithms for smaller and larger constraints. For a given
ensemble we can expect, in general, that the amount of work done (in terms of
workflows completed) depends mainly on the budget constraint, which determines
the number of machine-hours that can be provisioned. The deadline, although
important in terms of scheduling, is assumed to be greater than the critical
path of every workflow. We can then have a trivial but unrealistic solution
where the deadline  is long enough to allow all the workflow tasks to be
executed sequentially: this eliminates all the overheads of parallel execution
and leads to the highest resource utilization. More interesting and more
realistic, however, are the cases where the deadline is shorter than the serial
schedule, since it forces the system to provision multiple VMs to execute
workflow tasks in parallel.

@@

The number of workflows in an ensemble depends on the particular application, but
we assume that ensemble sizes are on the order of between 10 and 100 workflows.
This range is motivated by several reasons: First, such sizes are typical of the
real applications we have examined. For example, the number of geographical 
sites of interest to the users of the CyberShake application is on the order 
of 100. Second, smaller ensembles consisting of just a few workflows can be 
aggregated into a single workflow, so there is no need to treat them as an
ensemble. Similarly, when the number of workflows is larger than 100, and each 
workflow has a large number of tasks, either the constraints are low enough to
prevent many of the workflows from running, or the problem of efficiently 
allocating them to the resources becomes similar to the bag-of-tasks problem, 
which is easier to solve efficiently.

@@

In order to evaluate the algorithms on a standard set of workflows, we used
ensembles constructed from the workflows available\footnote{\url{https://confluence.pegasus.isi.edu/display/pegasus/WorkflowGenerator}}
in the workflow generator gallery~\cite{Bharathi08}. The gallery contains synthetic
workflows modeled on real applications. We used workflows from five applications 
that were managed by the Pegasus Workflow Management System. For each
application we constructed ensembles of 100 workflows by sampling collections of
workflows of various sizes (number of tasks) based on a given distribution.
Two distributions are considered:

@@

The constant distribution produces ensembles containing of workflows with the same number of tasks per workflow, 
while the Pareto-like distribution produces ensembles with a small number of larger 
workflows and a large number of smaller workflows, as seen in 
Fig.~\ref{fig:ensemble-distribution}. The number of large workflows (of size $\geq$ 900) 
is slightly increased to represent the ``heavy-tail'', i.e. that there is a 
not-insignificant number of large workflows in the ensemble. We assume that large
workflows are more important to users, so in the Pareto ensembles the workflows are 
assigned priorities according to their size so that the largest workflows are given 
the highest priority. In future work we plan to explore other ensemble types such as
uniform distributions.

@@

In the experiments we assumed that all the VMs are homogeneous, have a
processing speed of 1000 MIPS (million instructions per second), and a price of
\$1 per VM-hour. We also assume that the sizes of tasks in the workflow gallery
are given in seconds measured on a 1000 MIPS machine. These values were chosen
to approximate realistic numbers, and also to simplify interpretation of
results. Moreover, the absolute values of these parameters do not affect the
relative comparison between the algorithms. In this study we do not take into
account the heterogeneity of the infrastructure since we assume that it is
always possible to select a VM type that has the best price to performance ratio
for a given application.

For each application, we selected ranges of constraints (deadline and budget) so
that it is possible to observe the possibly broad result space: from tight 
constraints, where only a small number of workflows can be completed, to 
more liberal constraints where all, or almost all, of the workflows can be 
completed. These parameters vary among the applications. Montage and CyberShake 
ensembles required budgets in the \$20 -- \$150 range and deadlines from 1 to
20 hours to be completed. The LIGO and SIPHT workflows have more long-running
tasks, so their ensembles required budgets up to \$1,500 and deadlines up to 
40 and 60 hours. The Epigenomics workflow is even more costly, requiring 
approximately \$12,000 and 1500 hours to complete a full ensemble. These
differences in total runtime and cost come from the fact that we used the
same workflow size distribution for all five applications, but
the sizes of tasks vary between applications.

@@

\begin{itemize}
  \item $N_c$: number of workflows completed,
  \item $s$: user preference score (utility function),
  \item $T_C$: total amount of computing time completed,
  \item $C_{eff}$: effective cost of computing per hour, which includes
  overheads due to cost of idle resources.
\end{itemize}

@@

\begin{equation}
\label{eq:score}
s = \sum_{workflows\ completed}{2^{-d}}
\end{equation}

@@

The total amount of computing time completed, $T_C$, is calculated as the sum 
of the runtimes of all tasks in all completed workflows. Because we assumed 
that the resource cost per hour is equal to \$1, the total computing time 
can be easily compared to the allocated budget $b$. Their ratio: 
$u = (T_C * \$1)/b$, is the resource utilization, and its inverse: 
$C_{eff} = b/T_C$, is the effective resource cost in dollars per hour. To
calculate $C_{eff}$ and $u$ we use the value of allocated budget $b$, not the
actual total cost, since we are interested in the scenarios when the budget is
limited and does not allow completing all the workflows from the ensemble.
